\textls[-0]{\thetitle}

### 2023-03-17 metatheory-condensed.tex
edit on what's new about our W-types

\begin{itemize}
[labelsep=*,leftmargin=1pc,itemsep=3pt]
\item 
Linkages model families, so they are like tuples composed of fields
(with field names represented by variable bindings).
But there is a twist: linkages support late binding.
Unlike dependent tuples where a later component is \emph{existentially}
quantified over the earlier ones, a linkage component is \emph{universally}
quantified over---and thus polymorphic to---the components preceding it.\Zhixuan{explain the intuition behind the (typing rule of) linkages, and try to connect it to the ``family'' in the earlier section/surface syntax}\EDJreply{I am not sure if more is needed because the section ``Family polymorphism via linkages'' is trying to do this. Maybe I should estabilish connection between Rule l/add with the compilation process of fam poly?}\YZreply{I don't think you should mention anything about compilation process here. The surface syntax? Maybe. But I prefer not changing this sentence; it's supposed to be an executive summary rather than a detailed explanation.}\EDJreply{Actually I current am not sure where to add more sentences and how to elaborate more. Except for mentioning compilation process I don't think I have any good idea. So I will just stop here.}

@@

W-types~\cite{martin1984intuitionistic} are a succinct way to model
inductive types in MLTT.
Together with the identity type, they can express a
whole host of inductive types~\cite{hugunin2021not}. Specifically, with the traditional W-type and boolean types, we can encode an inductive type with multiple constructors.   
%
However, we formulate our W-type by an extra W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}) directly indicating each constructor and avoiding formentioned encoding. This is necessary to make our meta-theory syntax formulation closer to the surface syntax about inductive type and recursion handling supported by the plugin.\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types". Since there are already boolean types and Σ types, Fin types can be defined. Using Fin, ordinary W types suffice to define trees with n constructors. So it might be clearer to just define 'enhanced W-types' as syntactic sugar.}\YZreply{I concur.}\EDJreply{This following sentence has been replied to Zhixuan and he doesn't has further response: 1. We have to have enhanced W-type at FMLTT, that's for sure. Otherwise we (a). are too far away from the surface syntax because W-type itself only has one constructor all the time (b) this suggested encoding or syntactic sugar, since it interact with other connective like linkage (the Wrec is the interaction), so Wrec part is mostly unchanged  (c) since they will look the same, the encoding itself will look like an extra equality judgement in the syntax... which is not helpful for the complexity at all ... and even more complex  \\
2. so to me the problem you guys really have is around the content in the appendix -- I think Zhixuan mean he doesn't think this is more expressive. But my point is, it is hard to really write out the above syntactic sugar, formally. I think the issue here is similar to other reviewer saying "syntax translation". It is very hard to write out the encoding of Fin already, and then encode it again into W-type. (Just ask my agda mechanization). Of course, some hand-waiving can be done. For example, the part Zhixuan says I think it okay to hand-wave. But the destruction part -- the part Wrec interact with this encoding, I don't believe handwaving is enough. \\
3. But how do I say this to the audience? I mean the main reason I didn't choose syntactic translation/syntax sugar is basically ``doing encoding is harder than directly working in ambient logic.'' }
%
Each pair models a constructor of the inductive type.
The $i$-th pair, projected from the signature~$\wsig$ using the forms
$\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and
$\goodTerm{\Gm}{\wsigproj{i}{2}{\wsig}}{\wsigproj{i}{1}{\wsig} \to \cU}$,
defines the $i$-th constructor.
That is, given two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
$\goodTerm{\Gm, x : \El{\wsigproj{i}{2}{\wsig}(t_1)} }{t_2}{\El{\wcode{\wsig}}}$,
we can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type (\ruleref{tm/wsup}).
$\wcode{\wsig}$ gives the code of the
W-type (\ruleref{tm/w}).

>>>>

\begin{itemize}
[labelsep=*,leftmargin=1pc,itemsep=3pt]
\item 
Linkages model families, so they are like tuples composed of fields
(with field names represented by variable bindings).
But there is a twist: linkages support late binding.
Unlike dependent tuples where a later component is \emph{existentially}
quantified over the earlier ones, a linkage component is \emph{universally}
quantified over---and thus polymorphic to---the components preceding it.%\Zhixuan{explain the intuition behind the (typing rule of) linkages, and try to connect it to the ``family'' in the earlier section/surface syntax}\EDJreply{I am not sure if more is needed because the section ``Family polymorphism via linkages'' is trying to do this. Maybe I should estabilish connection between Rule l/add with the compilation process of fam poly?}\YZreply{I don't think you should mention anything about compilation process here. The surface syntax? Maybe. But I prefer not changing this sentence; it's supposed to be an executive summary rather than a detailed explanation.}\EDJreply{Actually I current am not sure where to add more sentences and how to elaborate more. Except for mentioning compilation process I don't think I have any good idea. So I will just stop here.}

@@

W-types~\cite{martin1984intuitionistic} are a succinct way to model
inductive types in MLTT.
Together with the identity type, they can express a whole host of inductive
types~\cite{hugunin2021not}, including those with multiple constructors.

Our formulation of W-types differs from previous ones in that it makes
individual constructors explicit and thus more closely matches the surface
syntax of \Lang.
%
In \TT, a W-type is given by a \emph{W-type signature} $\goodWSig{\Gm}{\wsig}{n}$ 
composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).
%\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types". Since there are already boolean types and Σ types, Fin types can be defined. Using Fin, ordinary W types suffice to define trees with n constructors. So it might be clearer to just define 'enhanced W-types' as syntactic sugar.}\YZreply{I concur.}\EDJreply{This following sentence has been replied to Zhixuan and he doesn't has further response: 1. We have to have enhanced W-type at FMLTT, that's for sure. Otherwise we (a). are too far away from the surface syntax because W-type itself only has one constructor all the time (b) this suggested encoding or syntactic sugar, since it interact with other connective like linkage (the Wrec is the interaction), so Wrec part is mostly unchanged  (c) since they will look the same, the encoding itself will look like an extra equality judgement in the syntax... which is not helpful for the complexity at all ... and even more complex  \\
%2. so to me the problem you guys really have is around the content in the appendix -- I think Zhixuan mean he doesn't think this is more expressive. But my point is, it is hard to really write out the above syntactic sugar, formally. I think the issue here is similar to other reviewer saying "syntax translation". It is very hard to write out the encoding of Fin already, and then encode it again into W-type. (Just ask my agda mechanization). Of course, some hand-waiving can be done. For example, the part Zhixuan says I think it okay to hand-wave. But the destruction part -- the part Wrec interact with this encoding, I don't believe handwaving is enough. \\
%3. But how do I say this to the audience? I mean the main reason I didn't choose syntactic translation/syntax sugar is basically ``doing encoding is harder than directly working in ambient logic.'' }
%
Each pair models a constructor of the inductive type.
The $i$-th pair, projected from the signature~$\wsig$ using the forms
$\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and
$\goodTerm{\Gm}{\wsigproj{i}{2}{\wsig}}{\wsigproj{i}{1}{\wsig} \to \cU}$,
defines the $i$-th constructor.
That is, given two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
$\goodTerm{\Gm, x : \El{\wsigproj{i}{2}{\wsig}(t_1)} }{t_2}{\El{\wcode{\wsig}}}$,
we can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type (\ruleref{tm/wsup}).
$\wcode{\wsig}$ gives the code of the
W-type (\ruleref{tm/w}).
